---
title: "Output-Based GVC Decomposition: The BM 2025 Framework"
author: "gvcAnalyzer Package"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Output-Based GVC Decomposition: The BM 2025 Framework}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
library(gvcAnalyzer)
```

# Introduction

This vignette provides a detailed examination of the **BM 2025 output-based GVC decomposition** methodology implemented in the **gvcAnalyzer** package. While traditional trade-based measures focus on exports, the output-based approach captures the full extent of GVC participation in a country's production structure.

## Theoretical Foundation

The BM 2025 framework (Borin & Mancini, 2025) decomposes gross output into five mutually exclusive components:

1. **Purely domestic production**: Value added that never crosses international borders
2. **Traditional trade**: Simple one-crossing exports absorbed directly abroad
3. **Pure-forward GVC**: Domestic value added that crosses borders multiple times (upstream suppliers)
4. **Pure-backward GVC**: Foreign value added embodied in production (downstream assemblers)
5. **Two-sided GVC**: Production involving both foreign inputs and subsequent re-export

This decomposition provides an output-side perspective complementary to trade-based measures.

## Reference

Borin, A., & Mancini, M. (2025). Measuring What Matters in Value-Added Trade: An Output-Based Approach. *Working Paper*.

# Setup and Data
```{r}
io <- bm_build_io(
  Z         = bm_toy_Z,
  Y         = bm_toy_Y,
  VA        = bm_toy_VA,
  X         = bm_toy_X,
  countries = bm_toy_countries,
  sectors   = bm_toy_sectors
)

io$countries
io$sectors
```

# Country-Level Output Decomposition

## Computing Output Components
```{r}
out_comp <- bm_2025_output_components(io)
out_comp
```

Each row represents a country's production structure decomposed into:

- **GVC_PF_X**: Pure-forward GVC-related output
- **GVC_PB_X**: Pure-backward GVC-related output
- **GVC_TSImp**: Two-sided GVC output from imported intermediates
- **GVC_TSDom**: Two-sided GVC output from domestic intermediates
- **GVC_TS_X**: Total two-sided GVC output
- **GVC_X**: Total GVC-related output
- **DomX**: Purely domestic output
- **TradX**: Traditional one-crossing trade
- **X_total**: Total gross output


At the country level, the decomposition satisfies the identity
`X_total = DomX + TradX + GVC_X`. At the world level, the sum of
`GVC_PF_X` equals the sum of `GVC_PB_X`, which provides a numerical
check on the balance between upstream and downstream GVC output.

## Example: China's Production Structure
```{r}
china_row <- out_comp[out_comp$country == "China", ]
china_row
```

## Output-Based Participation Measures
```{r}
out_meas <- bm_2025_output_measures(io)
out_meas
```

### Key Indicators

- **share_GVC_output**: Overall GVC participation intensity (GVC_X / X_total)
- **share_PF_output**: Share of GVC output in upstream activities (GVC_PF_X / GVC_X)
- **share_TS_output**: Share of GVC output in two-sided chains (GVC_TS_X / GVC_X)
- **share_PB_output**: Share of GVC output in downstream assembly (GVC_PB_X / GVC_X)
- **forward_output**: Forward orientation index ((GVC_PF_X − GVC_PB_X) / GVC_X)

A positive **forward_output** indicates upstream positioning, while negative values indicate downstream positioning.

### Comparing China and ROW
```{r}
china_meas <- out_meas[out_meas$country == "China", ]
row_meas   <- out_meas[out_meas$country == "ROW", ]

china_meas[, c("country", "share_GVC_output", "forward_output")]
row_meas[, c("country", "share_GVC_output", "forward_output")]
```

# Sector-Level Output Decomposition

## Computing Sectoral Components
```{r}
out_comp_sec <- bm_2025_output_components_sector(io)
head(out_comp_sec, 9)
```

This allocation ensures sectoral components aggregate to country totals.

### Verifying Aggregation Consistency
```{r}
check_agg <- aggregate(
  cbind(DomX_i, TradX_i, GVC_PF_Xi, GVC_PB_Xi, GVC_TS_Xi, GVC_Xi, X_i) ~ country,
  data = out_comp_sec,
  FUN = sum
)

check_agg
out_comp[, c("country", "DomX", "TradX", "GVC_PF_X", "GVC_PB_X", "GVC_TS_X", "GVC_X", "X_total")]
```

## Sectoral Participation Measures
```{r}
out_meas_sec <- bm_2025_output_measures_sector(io)
head(out_meas_sec, 9)
```

## Analyzing Sectoral Heterogeneity

### Ranking Sectors by GVC Intensity Within China
```{r}
china_sec <- out_meas_sec[out_meas_sec$country == "China", ]
china_sec <- china_sec[order(china_sec$share_GVC_output_i, decreasing = TRUE), ]
china_sec[, c("country", "sector", "share_GVC_output_i", "forward_output_i")]
```

### Cross-Country Comparison: Manufacturing Sector
```{r}
manuf_sec <- out_meas_sec[out_meas_sec$sector == "Manufacturing", ]
manuf_sec <- manuf_sec[order(manuf_sec$share_GVC_output_i, decreasing = TRUE), ]
manuf_sec[, c("country", "sector", "share_GVC_output_i", "forward_output_i")]
```

# Visualization

## Sectoral GVC Participation by Country
```{r fig.height=6, fig.width=7}
par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))
for (ctry in io$countries) {
  ctry_data <- out_meas_sec[out_meas_sec$country == ctry, ]
  barplot(
    height = ctry_data$share_GVC_output_i,
    names.arg = ctry_data$sector,
    main = paste(ctry, "GVC Participation"),
    ylab = "GVC Share of Output",
    col = "steelblue",
    las = 2
  )
}
```

## Forward vs Backward Orientation
```{r fig.width=7, fig.height=6}
par(mar = c(5, 5, 3, 2))
plot(
  out_meas_sec$share_PB_output_i,
  out_meas_sec$share_PF_output_i,
  pch = 19,
  col = as.numeric(factor(out_meas_sec$country)),
  cex = 1.5,
  xlab = "Backward GVC Share",
  ylab = "Forward GVC Share",
  main = "Sectoral GVC Orientation"
)
legend("topright", legend = io$countries, col = 1:io$G, pch = 19, cex = 0.8)
abline(a = 0, b = 1, lty = 2, col = "gray")
```

Sectors above the diagonal are more forward-oriented (upstream suppliers), while those below are more backward-oriented (downstream assemblers).

## Two-Sided GVC Decomposition
```{r}
ts_data <- out_comp_sec[out_comp_sec$GVC_TS_Xi > 0, ]
ts_data$TS_import_share <- ts_data$GVC_TSImp_i / ts_data$GVC_TS_Xi

head(ts_data[, c("country", "sector", "GVC_TS_Xi", "GVC_TSImp_i", 
                 "GVC_TSDom_i", "TS_import_share")], 9)
```

A higher **TS_import_share** indicates greater reliance on foreign inputs in two-sided GVC production.

# Applications

## Policy Analysis

The BM 2025 output decomposition is useful for:

1. **Industrial policy**: Identifying sectors with high GVC participation
2. **Supply chain resilience**: Assessing import dependence (GVC_TSImp)
3. **Upgrading strategies**: Tracking changes in forward/backward orientation
4. **Structural transformation**: Monitoring sectoral GVC role evolution

## Complementarity with Trade Measures

Output-based measures capture aspects not visible in trade statistics:

- Production for domestic final use embedding imported intermediates
- Domestic value added in goods exported by partners
- Total engagement in fragmented production beyond export intensity

# Summary

This vignette demonstrated:

1. Theoretical foundation of the BM 2025 output-based GVC framework
2. Country-level decomposition into five mutually exclusive output components
3. Participation measures including GVC intensity and forward/backward orientation
4. Sector-level decomposition preserving aggregation consistency
5. Sectoral heterogeneity in GVC participation and positioning
6. Visualization techniques for analyzing output-based GVC patterns
7. Two-sided GVC decomposition into imported versus domestic intermediate contributions

The output-based approach provides a comprehensive view of GVC participation complementary to traditional trade-based measures.

# Session Information
```{r}
sessionInfo()
```
